Bullet Data Preparation for LIME Diagnostics Paper

### 2020-09-25 code/02-data-preparation.Rmd
updated code to check for lands in one data but not the other (see section "Bullet Training Data", sub-section "Comparing Training Datasets", sub-sub-section "Land ID Differences" for the results)

@@

>>>>

## Comparing Training Datasets

We are trying to determine which of two datasets were used to train `rtrees` in `bulletxtrctr`.

@@

Load the newest training data provided by Heike (on 9/24/2020 from CSAFE data base):

```{r}
ccfs_withlands_raw = read.csv("../data/raw/CCFs_withlands.csv")
```

### Initial Comparisons

Data dimensions:

```{r}
dim(bullet_train_raw)
dim(ccfs_withlands_raw)
```

Comparing the distributions of the two datasets:

```{r}
summary(bullet_train_raw %>% filter(flag == FALSE) %>% select(rf_features))
summary(ccfs_withlands_raw %>% select(rf_features))
```

### Clean CCFs with lands

Rename some of the variables to match the structure of `bullet_train_raw`, select only variables of interest, and convert LEA labels to factors (for plotting purposes):

```{r}
ccfs_withlands_raw_expanded <- 
  ccfs_withlands_raw %>%
  rename("study1" = "study.x", "barrel1" = "barrel.x", "bullet1" = "bullet.x", "land1" = "land.x",
         "study2" = "study.y", "barrel2" = "barrel.y", "bullet2" = "bullet.y", "land2" = "land.y") %>%
  select(
    study1, barrel1, bullet1, land1, 
    study2, barrel2, bullet2, land2,
    all_of(rf_features), 
    match
  ) %>%
  mutate(across(c(study1, barrel1, bullet1, land1, study2, barrel2, bullet2, land2), factor))
```

Create a version of the `bullet_train_raw` data with the land IDs expanded to multiple columns for plotting:

```{r}
bullet_train_raw_expanded <- 
  bullet_train_raw %>%
  separate(land_id1, c("study1", "barrel1", "bullet1", "land1")) %>%
  separate(land_id2, c("study2", "barrel2", "bullet2", "land2")) %>%
  mutate(across(where(is.character), factor)) %>%
  mutate(barrel1 = fct_relevel(barrel1, paste0("Br", 1:10)),
         barrel2 = fct_relevel(barrel2, paste0("Br", 1:10)))
```

Shows that `Hamby44` needs to be changed to `Hamby173` in `ccfs_withlands`:

```{r fig.width = 12, fig.height = 5, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(study1, study2) %>%
    ggplot(aes(
      x = study1, y = study2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_raw_expanded %>% count(study1, study2) %>%
    ggplot(aes(
      x = study1, y = study2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

Shows that lettered barrels need to be changed to `BrUnk` and `Br` added to other barrels in `ccfs_withlands`:

```{r fig.width = 12, fig.height = 5, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(barrel1, barrel2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_raw_expanded %>% count(barrel1, barrel2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

Shows that barrel letters from in `ccfs_withlands` need to be made the bullet and `B` added to all bullets:

```{r fig.width = 12, fig.height = 5, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(bullet1, bullet2) %>%
    ggplot(aes(
      x = bullet1, y = bullet2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_raw_expanded %>% count(bullet1, bullet2) %>%
    ggplot(aes(
      x = bullet1, y = bullet2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

Shows that `L` needs to be added to lands in `ccfs_withlands`:

```{r fig.width = 12, fig.height = 5, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_raw_expanded %>% count(land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

Determine the letters in barrels:

```{r}
all_barrel_labels <-
  unique(c(
    as.character(ccfs_withlands_raw_expanded$barrel1),
    as.character(ccfs_withlands_raw_expanded$barrel2)
  ))
letters <- all_barrel_labels[!(all_barrel_labels %in% 1:10)]
letters
```

Clean `ccfs_withlands` so the labels match `bullet_train_raw`:

```{r}
ccfs_withlands_clean_expanded <- ccfs_withlands_raw_expanded %>%
  mutate(
    study1 = fct_recode(study1, "Hamby173" = "Hamby44"),
    study2 = fct_recode(study2, "Hamby173" = "Hamby44"),
    bullet1 = ifelse(barrel1 %in% letters, as.character(barrel1), as.character(bullet1)),
    barrel1 = ifelse(barrel1 %in% letters, "Unk", barrel1),
    bullet2 = ifelse(barrel2 %in% letters, as.character(barrel2), as.character(bullet2)),
    barrel2 = ifelse(barrel2 %in% letters, "Unk", barrel2)
    ) %>%
  mutate(
    study1 = fct_relevel(study1, "Hamby173", "Hamby252"),
    study2 = fct_relevel(study2, "Hamby173", "Hamby252"),
    barrel1 = paste0("Br", barrel1),
    bullet1 = paste0("B", bullet1),
    land1 = paste0("L", land1),
    barrel2 = paste0("Br", barrel2),
    bullet2 = paste0("B", bullet2),
    land2 = paste0("L", land2)
  ) %>% 
  mutate(across(where(is.character), factor)) %>%
  mutate(barrel1 = fct_relevel(barrel1, paste0("Br", 1:10)),
         barrel2 = fct_relevel(barrel2, paste0("Br", 1:10)))
```

The labels agree now but we are starting to see some differences in the two datasets:

```{r fig.width = 12, fig.height = 5, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(study1, study2) %>%
    ggplot(aes(
      x = study1, y = study2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_clean_expanded %>% count(study1, study2) %>%
    ggplot(aes(
      x = study1, y = study2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

The labels agree now but still can see the differences in the two datasets:

```{r fig.width = 12, fig.height = 5, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% count(barrel1, barrel2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_clean_expanded %>% count(barrel1, barrel2) %>%
    ggplot(aes(
      x = barrel1, y = barrel2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

The labels agree now but again, there are apparent differences between datasets:

```{r fig.width = 12, fig.height = 5, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% 
    count(bullet1, bullet2) %>%
    ggplot(aes(
      x = bullet1, y = bullet2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_clean_expanded %>% 
    count(bullet1, bullet2) %>%
    ggplot(aes(
      x = bullet1, y = bullet2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

The labels agree now:

```{r fig.width = 12, fig.height = 5, echo = FALSE}
plot_grid(
  bullet_train_raw_expanded %>% 
    count(land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) + 
    labs(title = "Hamby comparisons data"),
  ccfs_withlands_clean_expanded %>% 
    count(land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    theme(aspect.ratio = 1) +
    labs(title = "CCFs with lands data")
)
```

Compress the land labels for `ccfs_withlands`:

```{r}
ccfs_withlands_clean <- 
  ccfs_withlands_clean_expanded %>%
  mutate(
    land_id1 = paste(study1, barrel1, bullet1, land1, sep = "-"),
    land_id2 = paste(study2, barrel2, bullet2, land2, sep = "-")
  ) %>%
  select(land_id1, land_id2, all_of(rf_features), match)
```

### Land ID Differences

Extract the unique land IDs from the two datasets:

```{r}
ccfs_withlands_ids = unique(c(ccfs_withlands_clean$land_id1, ccfs_withlands_clean$land_id2))
bullet_train_ids = unique(c(bullet_train_raw$land_id1, bullet_train_raw$land_id1))
```

Identify the land ID in `CCFs_withlands` but not in `hamby-comparisons`:

```{r}
ccfs_withlands_ids[!(ccfs_withlands_ids %in% bullet_train_ids)]
```

Identify the land ID in `hamby-comparisons` but not in `CCFs_withlands`:

```{r}
bullet_train_ids[!(bullet_train_ids %in% ccfs_withlands_ids)]
```
